---
title: "Data Wrangling Assignment"
output: github_document
---

Time to learn about tibbles

```{r}
library(tidyverse)
#make a data frame into a tibble
as_tibble(iris)
tibble(
  x = 1:5,
  y = 1,
  z = x ^2 +y
)
```

tibbles can have invalid column names
use backticks for these ` ` also used in ggplot2, dplyr, and tidyr
```{r}
tb <- tibble(
  `)` = "smile",
  ` ` = "space",
  `2000` = "number"
)
tb

#tribble used to make tibbles
tribble(
  ~x, ~y, ~z,
  #--|--|----
  "a", 2, 3.6,
  "b", 1, 8.5
)
#how to print more and control it. n is rows.

nycflights13::flights %>%
  print(n = 10, width = Inf)

#or even better
nycflights13::flights %>%
  View()

print(mtcars)
```
tibbles show size, header information, type and the first 10 rows with as many columns as fit on screen.

use $ and [[ to extract information. $ will extract by name, and [[ will by position
```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)
df$x
df[["x"]]
df[[1]]
```
 to use those in a pipe use a period before . %>% .$x
 
 10.5 Exercises
 
 1. One is labled tbl_ the other data.frame when run. dataframe also prints all columns
 2. Data frames use an inexact searching method while tibbles are exact
```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[,"xyz"]
df[, c("abc", "xyz")]

tbl <- as_tibble(df)
tbl$x
tbl[, "xyz"]
tbl[, c("abc", "xyz")]
```
 
 3. [["mpg"]]
 4. 
```{r}
thing <- tibble(
  `1` = 1:5,
  `2` = 1,
  z = `1` ^2 +`2`
)
thing[["1"]]

ggplot(data = thing, aes(x =`1`, y = `2`)) + geom_point()
thing2 <- mutate(thing, `3` = `2` / `1`)
thing2 <- rename(thing, one = `1`, two = `2`)

```
     
5. tiblle::enframe() changes a vector name into a data frame with names and values
6. ?print.tbl, n_extra changes extra numbers columns.












read_csv() comma delimted
read_csv2() semicolon delimited
read_tsv() tab
read_delim() any delimiter???

read_fwf() fixed width files
  fwf_widths() or positions fwf_positions(). read_table columns separated by white space

read_log() reads APACHE LOGS! webreadr might be useful

```{r}
#heights <- read_csv("data/heights.csv")

```

```{r}
read_csv("a,b,c
1,2,3
4,5,6")

read_csv("The first line of metadata
  The second line of metadata
  x,y,z
  1,2,3", skip = 2)

# skip is useful for not printing out any metadata that may be on top # = lines skipped alternatively input # to skip all lines containing #


read_csv("# A comment I want to skip
  x,y,z
  1,2,3", comment = "#")

#if the first row does not have column names set col_names to false

read_csv("1,2,3\n4,5,6", col_names = FALSE)

#define the column header names to: x y z
read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))

#define what is the missing value value
read_csv("a,b,c\n1,2,.", na = ".")
```

Exercises
1. I would use read_delim(path, delim ="|")
2. col_names, col_types, locale, na, quoted_na, trim_ws, n_max, guess_max, progress
3. Path name, and col_positions
4. use read delim with quote = "'" and delimiter option as ","

```{r}
read_csv("a,b\n1,2,3\n4,5,6")
read_csv("a,b,c\n1,2\n1,2,3,4")

read_csv("a,b\n1,2\na,b")
read_csv("a;b\n1;3")
```

5. a mismatch of column headers and number of columns in some of the rows
  b miss match of number of columns and data
  c Open quote, and type mismatch
  d looks fine
  e semicolon is the delimiter, use csv2 instead

```{r}







#Super useful parser takes data and get it in a useful form

str(parse_logical(c("TRUE", "FALSE", "NA")))

str(parse_integer(c("1", "2", "3")))

str(parse_date(c("2010-01-01", "1979-10-14")))

parse_integer(c("1", "231", ".", "456"), na = ".")

x <- parse_integer(c("123", "345", "abc", "123.45"))

x

problems(x)

parse_double("1.23")

parse_double("1,23", locale = locale(decimal_mark = ","))
# will ignore non number characters
parse_number("$100")
#> [1] 100
parse_number("20%")
#> [1] 20
parse_number("It cost $123.45")
#> [1] 123

# Used in America
parse_number("$123,456,789")
#> [1] 1.23e+08

# Used in many parts of Europe
parse_number("123.456.789", locale = locale(grouping_mark = "."))
#> [1] 1.23e+08

# Used in Switzerland
parse_number("123'456'789", locale = locale(grouping_mark = "'"))
#> [1] 1.23e+08

charToRaw("Hadley")

x1 <- "El Ni\xf1o was particularly bad this year"
x2 <- "\x82\xb1\x82\xf1\x82\xc9\x82\xbf\x82\xcd"

x1
#> [1] "El Ni\xf1o was particularly bad this year"
x2
#> [1] "\x82\xb1\x82\xf1\x82ɂ\xbf\x82\xcd"

parse_character(x1, locale = locale(encoding = "Latin1"))
#> [1] "El Niño was particularly bad this year"
parse_character(x2, locale = locale(encoding = "Shift-JIS"))
#> [1] "こんにちは"

guess_encoding(charToRaw(x1))
#> # A tibble: 2 x 2
#>   encoding   confidence
#>   <chr>           <dbl>
#> 1 ISO-8859-1       0.46
#> 2 ISO-8859-9       0.23
guess_encoding(charToRaw(x2))
#> # A tibble: 1 x 2
#>   encoding confidence
#>   <chr>         <dbl>
#> 1 KOI8-R         0.42


fruit <- c("apple", "banana")
parse_factor(c("apple", "banana", "bananana"), levels = fruit)
#> Warning: 1 parsing failure.
#> row col           expected   actual
#>   3  -- value in level set bananana
#> [1] apple  banana <NA>  
#> attr(,"problems")
#> # A tibble: 1 x 4
#>     row   col expected           actual  
#>   <int> <int> <chr>              <chr>   
#> 1     3    NA value in level set bananana
#> Levels: apple banana

parse_datetime("2010-10-01T2010")
#> [1] "2010-10-01 20:10:00 UTC"
# If time is omitted, it will be set to midnight
parse_datetime("20101010")
#> [1] "2010-10-10 UTC"



parse_date("2010-10-01")
#> [1] "2010-10-01"

library(hms)
parse_time("01:10 am")
#> 01:10:00
parse_time("20:10:01")
#> 20:10:01

parse_date("01/02/15", "%m/%d/%y")
#> [1] "2015-01-02"
parse_date("01/02/15", "%d/%m/%y")
#> [1] "2015-02-01"
parse_date("01/02/15", "%y/%m/%d")
#> [1] "2001-02-15"

parse_date("1 janvier 2015", "%d %B %Y", locale = locale("fr"))
#> [1] "2015-01-01"

guess_parser("2010-10-01")
#> [1] "date"
guess_parser("15:01")
#> [1] "time"
guess_parser(c("TRUE", "FALSE"))
#> [1] "logical"
guess_parser(c("1", "5", "9"))
#> [1] "double"
guess_parser(c("12,352,561"))
#> [1] "number"

str(parse_guess("2010-10-10"))
#>  Date[1:1], format: "2010-10-10"

```
Exercises 
1. It has many arguments date_names, date_format, time_format, decimal_mark, grouping_mark, tz, encoding, asciify
2. You get an error. Period and comma switch! 
3. used for setting date and time formats

5. comma , or semicolon
6. 
```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14"
t1 <- "1705"
t2 <- "11:15:10.12 PM"

parse_date(d1, format = "%B %d, %Y")
parse_date(d2, "%Y-%b-%d")
parse_date(d3, "%d-%b-%Y")
parse_date(d4, "%B %d (%Y)")
parse_date(d5, "%m/%d/%y")
parse_time(t1, "%H%M")
parse_time(t2, "%I:%M:%OS %p")
```


```{r}





challenge <- read_csv(readr_example("challenge.csv"))

problems(challenge)

challenge <- read_csv(
  readr_example("challenge.csv"), 
  col_types = cols(
    x = col_integer(),
    y = col_character()
  )
)

challenge <- read_csv(
  readr_example("challenge.csv"), 
  col_types = cols(
    x = col_double(),
    y = col_character()
  )
)

tail(challenge)


challenge <- read_csv(
  readr_example("challenge.csv"), 
  col_types = cols(
    x = col_double(),
    y = col_date()
  )
)
tail(challenge)

challenge2 <- read_csv(readr_example("challenge.csv"), guess_max = 1001)

challenge2

challenge2 <- read_csv(readr_example("challenge.csv"), 
  col_types = cols(.default = col_character())
)

df <- tribble(
  ~x,  ~y,
  "1", "1.21",
  "2", "2.32",
  "3", "4.56"
)
df

type_convert(df)

write_csv(challenge, "challenge.csv")

challenge

write_csv(challenge, "challenge-2.csv")
read_csv("challenge-2.csv")

write_rds(challenge, "challenge.rds")
read_rds("challenge.rds")

library(feather)
write_feather(challenge, "challenge.feather")
read_feather("challenge.feather")


```


